Memory-efficient Mixed-Precision Implementations for\\Robust Explicit MPC

### 2019-04-08 emsoft-2019/Motiv.tex

	\includegraphics[width=3cm,height=5cm]{Figs/inv_pend.png}

@@

	\caption{Inverted pendulum}

@@

In this section, we use a classical control example to explain the motivation behind this work. Consider the problem of designing stablizing control for an inverted pendulum which takes (polytopic) constraints over state variables and control inputs into account. The linearized descrete-time description for the inverted pendulum depicted in Fig. \ref{fig:inverted_pendulum} is as follows:

@@

	where, $\theta$, $\omega$ and $u=\tau$ denote angular position, angular speed and input torque respectively; $g=9.81 m/s^2$ is the gravitational acceleration, $m$ the ball mass, $b$ the rotational fraction coefficient, $L$ the length of the bar and $T_s$ an appropriate sample time. Starting from an initial state, we would like the system's trajectory to converge into its equilibrium point while always $\theta\in[-\pi,\pi]$ and $\omega\in[-\pi/8,\pi/8]$. \\

@@

	However, one major challenge for implementing model predictive controllers over embedded systems will be to solve MPC optimality problem within each time step. Since the system under consideration is LTI, one can use explicit MPC scheme proposed in [Bemporad:2002] which computes partitions over state space together with affine functions which are used to compute the optimal control input over their corresponding partition. For the inverted-pendulum example, the computed controller contains $12$, over each an affine function is provided for generating the control input.\\

@@

	Overlay, we would like to design a robust explicit model predictive controller, taking errors resulted by fixed-point arithmetic into account and to come up with a minimal set of mixed precision assignments for all the coefficients that must be stored over the embedded system. Let us denote the total number of bits required for implementation of RMPC by $N_b$. Given the maximum memory budget denoted by $\bar N_b$, the overal goal will be to design a robust EMPC that (i) achieves performance objectives such as stability, safety and so on; (ii) it is implemented in a way that $N_b\leq \bar N_b$ while the error resulted by fixed-point arithmetic is smaller than the error bound $\Delta$. To that end, our approach is to start with an initial $\Delta$; design a robust explicit MPC controller with respect to $\Delta$ and to use state of the art fixed-point error analyzers to come up with the minimum total number of bits for storing the controller into the embedded system.\\

	Figure \ref{fig:overview} gives a high-level overview of our proposed setup. We start with definint the memory budget $\bar N_b$ and an initial error bound $\Delta_0$. Again, back to the inverted pendulum example, we choose $\bar N_b=6000$ and $\Delta_0=0$. To design robust explicit MPC for the error bound $\Delta$, MATLAB is run and the output controller is computed. Next, we ask Daisy to provide us with a mixed precision scheme respecting the error bound $\Delta=0$. However, $\Delta=0$ is not realizable as error resulted from fixed-point implementation is always greater than zero. Therefore, we go ahead and choose $\Delta=0.05$. Solving robust EMPC problem for the new disturbance bound gives the new controller. Next, we ask Daisy to provide us with fixed-point precision for all the parameters such that $\Delta=0.05$ is respected. The mixed precision scheme given by Daisy results in total number of bits $N_b=6084$ and since $N_b>\bar N_b$, we increase $\Delta$ again. Choosing $\Delta=0.1$, Daisy provides us with a mixed precision scheme with $N_b=5900$ for the correponding robust EMPC. As $N_b<\bar N_b$ the design goals are satisfied and the loop stops here. Surprisingly, the mixed precision implementation takes $11\%$ less memory comparing to the smallest uniform precision that respects $\Delta$. 
%	Using the output of Daisy, we are able to reduce memory requirements for the embedded systems up-to $20\%$ comparing to the case that uniform fixed-point precision is used.

>>>>

@@

	\begin{tabular}{cc}
	\includegraphics[width=2cm,height=4cm]{Figs/inv_pend.png}&	\includegraphics[width=6cm,height=4cm]{Figs/regs.jpg}\\
	(a)&(b)
	\end{tabular}
	\caption{(a) Inverted pendulum; (b) 2-D plot of explicit MPC polyhedral partitions for the inverted pendulum}

@@

In this section, we use a classical control example to explain the motivation behind this work. Consider the problem of designing stablizing control for an inverted pendulum which takes (polytopic) constraints over state variables and control inputs into account. The linearized discrete-time description for the inverted pendulum depicted in Fig. \ref{fig:inverted_pendulum} (a) is as follows:

@@

	where, $\theta$, $\omega$ and $u=\tau$ denote angular position, angular speed and input torque respectively; $g=9.81 m/s^2$ is the gravitational acceleration, $m$ the ball mass, $b$ the rotational fraction coefficient, $L$ the length of the bar and $T_s$ an appropriate sample time. Starting from an initial state, we would like the system's trajectory to converge into its equilibrium point ($\theta=0$ and $\omega=0$) while $\theta\in[-\pi,\pi]$ and $\omega\in[-\pi/8,\pi/8]$ within the time horizon $N=2$. \\

@@

	However, one major challenge for implementing model predictive controllers over embedded systems will be to solve MPC optimality problem within each time step. Since the system under consideration is LTI, one can use explicit MPC scheme proposed in \cite{Bemporad:2002} which computes partitions over state space together with affine functions which are used to compute the optimal control input over their corresponding partition. For the inverted-pendulum example, the computed controller contains $12$, over each an affine function is provided for generating the control input.\\

@@

	Overlay, we would like to design a robust explicit model predictive controller, taking errors resulted by fixed-point arithmetic into account and to come up with a minimal set of mixed precision assignments for all the coefficients that must be stored over the embedded system. Let us denote the total number of bits required for implementation of RMPC by $N_b$. Given the maximum memory budget denoted by $\bar N_b$, the overall goal will be to design a robust EMPC that (i) achieves performance objectives such as stability, safety and so on; (ii) it is implemented in a way that $N_b\leq \bar N_b$ while the error resulted by fixed-point arithmetic is smaller than the error bound $\Delta$. To that end, our approach is to start with an initial $\Delta$; design a robust explicit MPC controller with respect to $\Delta$ and to use state of the art fixed-point error analyzers to come up with the minimum total number of bits for storing the controller into the embedded system.\\

	Figure \ref{fig:overview} gives a high-level overview of our proposed setup. We start with defining the memory budget $\bar N_b$ and an initial error bound $\Delta_0$. Again, back to the inverted pendulum example, we choose $\bar N_b=6000$ and $\Delta_0=0$. To design robust explicit MPC for the error bound $\Delta$, MATLAB is run and the output controller is computed. Next, we ask Daisy to provide us with a mixed precision scheme respecting the error bound $\Delta=0$. However, $\Delta=0$ is not realizable as error resulted from fixed-point implementation is always greater than zero. Therefore, we go ahead and choose $\Delta=0.05$. Solving robust EMPC problem for the new disturbance bound gives the new controller. Next, we ask Daisy to provide us with fixed-point precision for all the parameters such that $\Delta=0.05$ is respected. The mixed precision scheme given by Daisy results in total number of bits $N_b=6084$ and since $N_b>\bar N_b$, we increase $\Delta$ again. Choosing $\Delta=0.1$, Daisy provides us with a mixed precision scheme with $N_b=5900$ for the corresponding robust EMPC. As $N_b<\bar N_b$ the design goals are satisfied and the loop stops here. Surprisingly, the mixed precision implementation takes $11\%$ less memory comparing to the smallest uniform precision that respects $\Delta$. 
%	Using the output of Daisy, we are able to reduce memory requirements for the embedded systems up-to $20\%$ comparing to the case that uniform fixed-point precision is used.
